---
title: "cazyme_diversity_ver1"
author: "Tord Ranheim"
date: "2024-10-28"
output:
  html_document:
    keep_md: no
---

CAZyme diversity


```{r setup, include=FALSE}
knitr::opts_chunk$set(dev = "png",
                      dpi=300,
                      echo=F,
                      cache=T, echo=T)
```



###loading requisite libraries 
```{r libraries to load}
library(tidyverse)
library(rmarkdown)
library(knitr)
library(magrittr)
library(ggplot2)
library(dplyr)
library(phyloseq)
library(phylosmith)
library(microbiome)
library(vegan)
library(metagMisc)
library(microViz)

```


##Clear the directory and set pathway
```{r}
rm(list=ls())

```


Data loading
```{r Load phyloseq objects}

#fungal cazymes
ps_fungi = readRDS("/Users/tranheim/Desktop/Studier/Remiil-NFI/Study 1/Processing/Microbial/Metagenome/Processing/CAZymes/Results/ps_caz_fun_rar.RDS")

#bacterial cazyme
ps_bacteria = readRDS("/Users/tranheim/Desktop/Studier/Remiil-NFI/Study 1/Processing/Microbial/Metagenome/Processing/CAZymes/Results/ps_caz_bac_rar.RDS")

```


```{r Load metadata file}

metadata = read.table("/Users/tranheim/Desktop/Studier/Remiil-NFI/Study 1/Processing/Metadata/Result/metadata.txt")
row.names(metadata) = metadata$Sample

```


```{r}

table(ps_bacteria@sam_data$Succession)

```


We are interested in comparing grasslands to their successional endpoint (i.e. Forests), but also to compare different grassland successional stages against each other. For this we create a new metadata set where the forest level is set as reference level for linear models.

```{r Set forest as reference level}

metadata_ref = metadata
metadata_ref$Succession <- as.factor(metadata_ref$Succession) #Set successional stage as factor
metadata_ref$Succession <- relevel(metadata_ref$Succession, ref="Forest") #Set forest as reference level

```


                                  
                                        ESTIMATE DIVERSITY
###################################################################################################

We estimate both shannon diversity (accounting for taxon richness and relative abundances), as well observed species richness

```{r Estimate diversities}

t1 = ps_fungi %>% estimate_richness(measures = "Shannon")
t2 = ps_bacteria %>% estimate_richness(measures = "Shannon")
names(t1) = "cazyme_diversity_fungi"
t1$Sample = row.names(t1)
names(t2) = "cazyme_diversity_bacteria"
t2$Sample = row.names(t2)

#Add to metadata
metadata = left_join(metadata, t1)
metadata = left_join(metadata, t2)
metadata_ref = left_join(metadata_ref, t1)
metadata_ref = left_join(metadata_ref, t2)


rm(t1,t2)
```



                                  TEST FOR DIFFERENCES - FUNGI
###################################################################################################
```{r Load libraries needed for LMMs}
library(caret)
library(DHARMa)
library(jtools)
library (multcomp)
library(lme4)
library(car)

```


Now we test for differences in diversity using linear mixed effect models incorporating the paired structure of the dataset as random factors, together with differences in parent material

```{r test for differences in fungal cazyme diversity}

#Check data distribution
hist(metadata$cazyme_diversity_fungi, xlab = "Response", main = "")
shapiro.test(metadata$cazyme_diversity_fungi) #Data is normally distributed

#Test for differences in grasslands relative to forest reference level
mod1 <- lmer(cazyme_diversity_fungi ~ Succession + (1|Par) + (1|Parent_material) + (1|Distance),
             data = metadata_ref)

#Check model disagnostics
resid <- simulateResiduals(fittedModel = mod1, plot = F)
plot(resid) #QQ-plot looks ok, residual plot shows some deviations but not much

#Inspect results
summary(mod1)
summ(mod1, exp = T)

#Allgrasslands differ from forests

#Now we test for differences between grassland successional stages using the same models but without forests as reference levels
mod1 <- lmer(cazyme_diversity_fungi ~ Succession +  (1|Par) + (1|Parent_material) + (1|Distance),
             data = metadata)
summary(mod1)
summ(mod1, exp = T)

#This is followed bymultiple comparisons between all stages
summary(glht(mod1, mcp(Succession  = "Tukey")))
#No differences between any of the grassland successional stages 


#Estimate effect size using eta squared
mod2 = lmer(cazyme_diversity_fungi ~ Ecosystem_type + (1|Par) + 
              (1|Parent_material) + (1|Distance), REML = T, data = metadata)

aov = Anova(mod2, type = 2, test = "F")
aov

d1 = effectsize::omega_squared(aov, alternative = "two.sided")
d1$Community = "Fungi"

```



                              TEST FOR DIFFERENCES - BACTERIA
###################################################################################################

```{r bacterial cazyme diversity}

#Check data distribution
hist(metadata$cazyme_diversity_bacteria, xlab = "Response", main = "")
shapiro.test(metadata$cazyme_diversity_bacteria) #Data is normally distributed

#Run model
mod1 <- lmer(cazyme_diversity_bacteria ~ Succession +  (1|Par) + (1|Parent_material) + (1|Distance),
             data = metadata_ref)

#Check model disagnostics
resid <- simulateResiduals(fittedModel = mod1, plot = F)
plot(resid) #QQ-plot looks ok, residual plots look ok

#Inspect results
summary(mod1)
summ(mod1, exp = T)


#Now we test for differences between grassland successional stages using the same models but without forests as reference levels
mod1 <- lmer(cazyme_diversity_bacteria ~ Succession +  (1|Par) + (1|Parent_material) + (1|Distance),
             data = metadata)
summary(mod1)
summ(mod1, exp = T)

#This is followed bymultiple comparisons between all stages
summary(glht(mod1, mcp(Succession  = "Tukey")))
#No differences between any of the grassland successional stages 


#Estimate effect size using eta squared
mod2 = lmer(cazyme_diversity_bacteria ~ Ecosystem_type + (1|Par),
              REML = T, data = metadata)

aov = Anova(mod2, type = 2, test = "F")
aov

d2 = effectsize::omega_squared(aov, alternative = "two.sided")
d2$Community = "Bacteria"

```


```{r Save diversity estimates for plotting}

t1 = metadata %>% dplyr::select(Sample, cazyme_diversity_fungi, cazyme_diversity_bacteria)
t2 = rbind(d1,d2)

write.table(t1, file = "./Results/cazyme_diversity.txt")
write.table(t2, file = "./Results/es_cazyme_diversity.txt")

```


