---
title: "sir_ver1"
author: "Tord Ranheim"
date: "2024-11-13"
output:
  html_document:
    keep_md: no
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(dev = "png",
                      dpi=300,
                      echo=F,
                      cache=T, echo=T)
```



###loading requisite libraries 
```{r libraries to load}
library(tidyverse)
library(rmarkdown)
library(knitr)
library(magrittr)
library(ggplot2)
library(RColorBrewer)
library(ggpubr)
library(ggsci)
library(dplyr)
library(phyloseq)
library(phylosmith)
library(microbiome)
library(vegan)
library(metagMisc)
library(ape)
library(microViz)

```


##Clear the directory and set pathway
```{r}
rm(list=ls())

```


```{r Load metadata file}

metadata = read.table("/Users/tranheim/Desktop/Studier/Remiil-NFI/Study 1/Processing/Metadata/Result/metadata.txt")
row.names(metadata) = metadata$Sample

env = read.table("/Users/tranheim/Desktop/Studier/Remiil-NFI/Study 1/Processing/Metadata/Partitioning/Results/env_imputed.txt")
env$Succession = metadata$Succession

```


                                        DATA PREPARATION
###################################################################################################

```{r Create subset of sir data}

sir = metadata %>% dplyr::select(Basal_resp, Glucose_resp, Glycine_resp,
                                 OA_resp, AY_resp, Lignin_resp, Chitin_resp)
  

sir$msir = rowSums(sir[,2:7]) #Add multiple substrate-induced respiration column

#Standardize each individual substrate respiration by msir
sir$Glucose_resp_std = sir$Glucose_resp/sir$msir
sir$Glycine_resp_std = sir$Glycine_resp/sir$msir
sir$OA_resp_std = sir$OA_resp/sir$msir
sir$AY_resp_std = sir$AY_resp/sir$msir
sir$Lignin_resp_std = sir$Lignin_resp/sir$msir
sir$Chitin_resp_std = sir$Chitin_resp/sir$msir

sir$Sample = row.names(sir)
env$Sample = row.names(env)
m1 = left_join(env, sir)
m1$Par = metadata$Par
m1$Distance = metadata$Distance

rm(sir, env)  

#Write
write.table(m1, file = "res_sir.txt")

```



We are interested in comparing grasslands to their successional endpoint (i.e. Forests), but also to compare different grassland successional stages against each other. For this we create a new metadata set where the forest level is set as reference level for linear models.

```{r Set forest as reference level}

m1$Succession <- as.factor(m1$Succession) #Set successional stage as factor
m1$Succession <- relevel(m1$Succession, ref="Forest") #Set forest as reference level

```

```{r Set plotting details}

m2 = m1

#Set order of plotting
m2$Succession = ordered(m2$Succession, levels = c("Managed","Recent","Late","Forest"))

#Set manual color and shape
farg = c("#bd0026","#fd8d3c","#fecc5c", "forestgreen")
shape = c(21,23,24,25)
farg2 = "grey59"

#Create median line for forest reference stage
f = m2 %>% filter(Succession == "Forest") 

```


                                  TEST FOR DIFFERENCES
###################################################################################################
```{r Load libraries needed for LMMs}
library(caret)
library(DHARMa)
library(jtools)
library (multcomp)
library(lme4)
library(MASS)

```

```{r basal resp}

h = median(f$Basal_resp)

m2 %>%
    bind_rows(tibble(var1 = "trt99")) %>% 
  ggplot(aes(x= Succession, y = Basal_resp, fill = Succession))+
  geom_boxplot(aes(fill= Succession),outlier.shape = NA)+
    geom_point(position=position_jitter(0.1, seed = 123),alpha=0.5)+
    scale_fill_manual(values=farg, na.translate = F)+
    scale_color_manual(values=farg, na.translate = F)+
      labs(fill = "Land use", y = "Diversity")+
      #ylim(1,6)+
      theme_classic()+
  #geom_line(aes(group = Par),alpha = 0.05)+
  scale_x_discrete(limits = c("Managed","Recent","Late","trt99","Forest"),
  breaks = c("Managed","Recent","Late",NA,"Forest"),
  labels = c("Managed","Recent","Late",NA,"Forest"))+
  geom_vline(xintercept = "trt99", linetype = 2, linewidth = 0.2)+
    geom_hline(yintercept = h, linetype = 2, linewidth = 0.9, col = "black")+
  theme(axis.title.x = element_blank())+
  theme(axis.text.x = element_blank())+
    theme(axis.title.y = element_blank())+
  theme(axis.ticks.x = element_blank())+
  #theme(legend.position = "none")+
     theme(text=element_text(size=20))


#Check data distribution
hist(m1$Basal_resp, xlab = "Response", main = "")
shapiro.test(m1$Basal_resp) #Data is skewed: use GLMMs

#Test for differences in grasslands relative to forest reference level: Gamma distribution
mod1 <- glm(Basal_resp ~ Succession + (1|Par)  + (1|Distance),
             family = Gamma, 
             data = m1)


#Check model disagnostics
resid <- simulateResiduals(fittedModel = mod1, plot = F)
plot(resid) #Ok

#Inspect results
summary(mod1)
summ(mod1, exp = T)

```
```{r Glucose}

#Glucose
m2 %>%
    bind_rows(tibble(var1 = "trt99")) %>% 
  ggplot(aes(x= Succession, y = Glucose_resp, fill = Succession))+
  geom_boxplot(aes(fill= Succession),outlier.shape = NA)+
    geom_point(position=position_jitter(0.1, seed = 123),alpha=0.5)+
    scale_fill_manual(values=farg, na.translate = F)+
    scale_color_manual(values=farg, na.translate = F)+
      labs(fill = "Land use", y = "Diversity")+
      #ylim(1,6)+
      theme_classic()+
  #geom_line(aes(group = Par),alpha = 0.05)+
  scale_x_discrete(limits = c("Managed","Recent","Late","trt99","Forest"),
  breaks = c("Managed","Recent","Late",NA,"Forest"),
  labels = c("Managed","Recent","Late",NA,"Forest"))+
  geom_vline(xintercept = "trt99", linetype = 2, linewidth = 0.2)+
    geom_hline(yintercept = h, linetype = 2, linewidth = 0.9, col = "black")+
  theme(axis.title.x = element_blank())+
  theme(axis.text.x = element_blank())+
    theme(axis.title.y = element_blank())+
  theme(axis.ticks.x = element_blank())+
  #theme(legend.position = "none")+
     theme(text=element_text(size=20))

#Check data distribution
hist(m1$Glucose_resp, xlab = "Response", main = "")

#Test for differences in grasslands relative to forest reference level: Gamma distribution
mod1 <- glm(Glucose_resp ~ Succession + (1|Par)  + (1|Distance),
             family = gaussian, 
             data = m1)

#Check model disagnostics
resid <- simulateResiduals(fittedModel = mod1, plot = F)
plot(resid) #Ok

#Inspect results
summary(mod1)
summ(mod1, exp = T)

```


```{r Glucose standardized}

#Glucose
m2 %>%
    bind_rows(tibble(var1 = "trt99")) %>% 
  ggplot(aes(x= Succession, y = Glucose_resp_std, fill = Succession))+
  geom_boxplot(aes(fill= Succession),outlier.shape = NA)+
    geom_point(position=position_jitter(0.1, seed = 123),alpha=0.5)+
    scale_fill_manual(values=farg, na.translate = F)+
    scale_color_manual(values=farg, na.translate = F)+
      labs(fill = "Land use", y = "Diversity")+
      #ylim(1,6)+
      theme_classic()+
  #geom_line(aes(group = Par),alpha = 0.05)+
  scale_x_discrete(limits = c("Managed","Recent","Late","trt99","Forest"),
  breaks = c("Managed","Recent","Late",NA,"Forest"),
  labels = c("Managed","Recent","Late",NA,"Forest"))+
  geom_vline(xintercept = "trt99", linetype = 2, linewidth = 0.2)+
    geom_hline(yintercept = h, linetype = 2, linewidth = 0.9, col = "black")+
  theme(axis.title.x = element_blank())+
  theme(axis.text.x = element_blank())+
    theme(axis.title.y = element_blank())+
  theme(axis.ticks.x = element_blank())+
  #theme(legend.position = "none")+
     theme(text=element_text(size=20))

#Check data distribution
hist(m1$Glucose_resp_std, xlab = "Response", main = "")

#Test for differences in grasslands relative to forest reference level: Gamma distribution
mod1 <- glm(Glucose_resp_std ~ Succession + (1|Par)  + (1|Distance),
             family = gaussian, 
             data = m1)

#Check model disagnostics
resid <- simulateResiduals(fittedModel = mod1, plot = F)
plot(resid) #Ok

#Inspect results
summary(mod1)
summ(mod1, exp = T)

```


```{r Oxalic acid}


m2 %>%
    bind_rows(tibble(var1 = "trt99")) %>% 
  ggplot(aes(x= Succession, y = OA_resp, fill = Succession))+
  geom_boxplot(aes(fill= Succession),outlier.shape = NA)+
    geom_point(position=position_jitter(0.1, seed = 123),alpha=0.5)+
    scale_fill_manual(values=farg, na.translate = F)+
    scale_color_manual(values=farg, na.translate = F)+
      labs(fill = "Land use", y = "Diversity")+
      #ylim(1,6)+
      theme_classic()+
  #geom_line(aes(group = Par),alpha = 0.05)+
  scale_x_discrete(limits = c("Managed","Recent","Late","trt99","Forest"),
  breaks = c("Managed","Recent","Late",NA,"Forest"),
  labels = c("Managed","Recent","Late",NA,"Forest"))+
  geom_vline(xintercept = "trt99", linetype = 2, linewidth = 0.2)+
    geom_hline(yintercept = h, linetype = 2, linewidth = 0.9, col = "black")+
  theme(axis.title.x = element_blank())+
  theme(axis.text.x = element_blank())+
    theme(axis.title.y = element_blank())+
  theme(axis.ticks.x = element_blank())+
  #theme(legend.position = "none")+
     theme(text=element_text(size=20))

#Check data distribution
hist(m1$OA_resp, xlab = "Response", main = "")

#Test for differences in grasslands relative to forest reference level: Gamma distribution
mod1 <- glm(OA_resp ~ Succession + (1|Par)  + (1|Distance),
             family = Gamma, 
             data = m1)

#Check model disagnostics
resid <- simulateResiduals(fittedModel = mod1, plot = F)
plot(resid) #Ok

#Inspect results
summary(mod1)
summ(mod1, exp = T)

```


```{r Oxalic acid standardized}

m2 %>%
    bind_rows(tibble(var1 = "trt99")) %>% 
  ggplot(aes(x= Succession, y = OA_resp_std, fill = Succession))+
  geom_boxplot(aes(fill= Succession),outlier.shape = NA)+
    geom_point(position=position_jitter(0.1, seed = 123),alpha=0.5)+
    scale_fill_manual(values=farg, na.translate = F)+
    scale_color_manual(values=farg, na.translate = F)+
      labs(fill = "Land use", y = "Diversity")+
      #ylim(1,6)+
      theme_classic()+
  #geom_line(aes(group = Par),alpha = 0.05)+
  scale_x_discrete(limits = c("Managed","Recent","Late","trt99","Forest"),
  breaks = c("Managed","Recent","Late",NA,"Forest"),
  labels = c("Managed","Recent","Late",NA,"Forest"))+
  geom_vline(xintercept = "trt99", linetype = 2, linewidth = 0.2)+
    geom_hline(yintercept = h, linetype = 2, linewidth = 0.9, col = "black")+
  theme(axis.title.x = element_blank())+
  theme(axis.text.x = element_blank())+
    theme(axis.title.y = element_blank())+
  theme(axis.ticks.x = element_blank())+
  #theme(legend.position = "none")+
     theme(text=element_text(size=20))

#Check data distribution
hist(m1$OA_resp_std, xlab = "Response", main = "")

#Test for differences in grasslands relative to forest reference level: Gamma distribution
mod1 <- glm(OA_resp_std ~ Succession + (1|Par)  + (1|Distance),
             family = gaussian, 
             data = m1)

#Check model disagnostics
resid <- simulateResiduals(fittedModel = mod1, plot = F)
plot(resid) #Ok

#Inspect results
summary(mod1)
summ(mod1, exp = T)

```


```{r Autolyzed yeast}

m2 %>%
    bind_rows(tibble(var1 = "trt99")) %>% 
  ggplot(aes(x= Succession, y = AY_resp, fill = Succession))+
  geom_boxplot(aes(fill= Succession),outlier.shape = NA)+
    geom_point(position=position_jitter(0.1, seed = 123),alpha=0.5)+
    scale_fill_manual(values=farg, na.translate = F)+
    scale_color_manual(values=farg, na.translate = F)+
      labs(fill = "Land use", y = "Diversity")+
      #ylim(1,6)+
      theme_classic()+
  #geom_line(aes(group = Par),alpha = 0.05)+
  scale_x_discrete(limits = c("Managed","Recent","Late","trt99","Forest"),
  breaks = c("Managed","Recent","Late",NA,"Forest"),
  labels = c("Managed","Recent","Late",NA,"Forest"))+
  geom_vline(xintercept = "trt99", linetype = 2, linewidth = 0.2)+
    geom_hline(yintercept = h, linetype = 2, linewidth = 0.9, col = "black")+
  theme(axis.title.x = element_blank())+
  theme(axis.text.x = element_blank())+
    theme(axis.title.y = element_blank())+
  theme(axis.ticks.x = element_blank())+
  #theme(legend.position = "none")+
     theme(text=element_text(size=20))

#Check data distribution
hist(m1$AY_resp, xlab = "Response", main = "")

#Test for differences in grasslands relative to forest reference level: Gamma distribution
mod1 <- glm(AY_resp ~ Succession + (1|Par)  + (1|Distance),
             family = Gamma, 
             data = m1)

#Check model disagnostics
resid <- simulateResiduals(fittedModel = mod1, plot = F)
plot(resid) #Ok

#Inspect results
summary(mod1)
summ(mod1, exp = T)

```

```{r Autolyzed yeast standardized}

m2 %>%
    bind_rows(tibble(var1 = "trt99")) %>% 
  ggplot(aes(x= Succession, y = AY_resp_std, fill = Succession))+
  geom_boxplot(aes(fill= Succession),outlier.shape = NA)+
    geom_point(position=position_jitter(0.1, seed = 123),alpha=0.5)+
    scale_fill_manual(values=farg, na.translate = F)+
    scale_color_manual(values=farg, na.translate = F)+
      labs(fill = "Land use", y = "Diversity")+
      #ylim(1,6)+
      theme_classic()+
  #geom_line(aes(group = Par),alpha = 0.05)+
  scale_x_discrete(limits = c("Managed","Recent","Late","trt99","Forest"),
  breaks = c("Managed","Recent","Late",NA,"Forest"),
  labels = c("Managed","Recent","Late",NA,"Forest"))+
  geom_vline(xintercept = "trt99", linetype = 2, linewidth = 0.2)+
    geom_hline(yintercept = h, linetype = 2, linewidth = 0.9, col = "black")+
  theme(axis.title.x = element_blank())+
  theme(axis.text.x = element_blank())+
    theme(axis.title.y = element_blank())+
  theme(axis.ticks.x = element_blank())+
  #theme(legend.position = "none")+
     theme(text=element_text(size=20))

#Check data distribution
hist(m1$AY_resp_std, xlab = "Response", main = "")

#Test for differences in grasslands relative to forest reference level: Gamma distribution
mod1 <- glm(AY_resp_std ~ Succession + (1|Par)  + (1|Distance),
             family = Gamma, 
             data = m1)

#Check model disagnostics
resid <- simulateResiduals(fittedModel = mod1, plot = F)
plot(resid) #Ok

#Inspect results
summary(mod1)
summ(mod1, exp = T)

```


```{r Lignin}

m2 %>%
    bind_rows(tibble(var1 = "trt99")) %>% 
  ggplot(aes(x= Succession, y = Lignin_resp, fill = Succession))+
  geom_boxplot(aes(fill= Succession),outlier.shape = NA)+
    geom_point(position=position_jitter(0.1, seed = 123),alpha=0.5)+
    scale_fill_manual(values=farg, na.translate = F)+
    scale_color_manual(values=farg, na.translate = F)+
      labs(fill = "Land use", y = "Diversity")+
      #ylim(1,6)+
      theme_classic()+
  #geom_line(aes(group = Par),alpha = 0.05)+
  scale_x_discrete(limits = c("Managed","Recent","Late","trt99","Forest"),
  breaks = c("Managed","Recent","Late",NA,"Forest"),
  labels = c("Managed","Recent","Late",NA,"Forest"))+
  geom_vline(xintercept = "trt99", linetype = 2, linewidth = 0.2)+
    geom_hline(yintercept = h, linetype = 2, linewidth = 0.9, col = "black")+
  theme(axis.title.x = element_blank())+
  theme(axis.text.x = element_blank())+
    theme(axis.title.y = element_blank())+
  theme(axis.ticks.x = element_blank())+
  #theme(legend.position = "none")+
     theme(text=element_text(size=20))

#Check data distribution
hist(m1$Lignin_resp, xlab = "Response", main = "")

#Test for differences in grasslands relative to forest reference level: Gamma distribution
mod1 <- glm(Lignin_resp ~ Succession + (1|Par)  + (1|Distance),
             family = gaussian, 
             data = m1)

#Check model disagnostics
resid <- simulateResiduals(fittedModel = mod1, plot = F)
plot(resid) #Ok

#Inspect results
summary(mod1)
summ(mod1, exp = T)

```


```{r Lignin standardized}

m2 %>%
    bind_rows(tibble(var1 = "trt99")) %>% 
  ggplot(aes(x= Succession, y = Lignin_resp_std, fill = Succession))+
  geom_boxplot(aes(fill= Succession),outlier.shape = NA)+
    geom_point(position=position_jitter(0.1, seed = 123),alpha=0.5)+
    scale_fill_manual(values=farg, na.translate = F)+
    scale_color_manual(values=farg, na.translate = F)+
      labs(fill = "Land use", y = "Diversity")+
      #ylim(1,6)+
      theme_classic()+
  #geom_line(aes(group = Par),alpha = 0.05)+
  scale_x_discrete(limits = c("Managed","Recent","Late","trt99","Forest"),
  breaks = c("Managed","Recent","Late",NA,"Forest"),
  labels = c("Managed","Recent","Late",NA,"Forest"))+
  geom_vline(xintercept = "trt99", linetype = 2, linewidth = 0.2)+
    geom_hline(yintercept = h, linetype = 2, linewidth = 0.9, col = "black")+
  theme(axis.title.x = element_blank())+
  theme(axis.text.x = element_blank())+
    theme(axis.title.y = element_blank())+
  theme(axis.ticks.x = element_blank())+
  #theme(legend.position = "none")+
     theme(text=element_text(size=20))

#Check data distribution
hist(m1$Lignin_resp_std, xlab = "Response", main = "")

#Test for differences in grasslands relative to forest reference level: Gamma distribution
mod1 <- glm(Lignin_resp_std ~ Succession + (1|Par)  + (1|Distance),
             family = gaussian, 
             data = m1)

#Check model disagnostics
resid <- simulateResiduals(fittedModel = mod1, plot = F)
plot(resid) #Ok

#Inspect results
summary(mod1)
summ(mod1, exp = T)

```


```{r Chitin standardized}

m2 %>%
  filter(Chitin_resp > 0) %>%
    bind_rows(tibble(var1 = "trt99")) %>% 
  ggplot(aes(x= Succession, y = Chitin_resp, fill = Succession))+
  geom_boxplot(aes(fill= Succession),outlier.shape = NA)+
    geom_point(position=position_jitter(0.1, seed = 123),alpha=0.5)+
    scale_fill_manual(values=farg, na.translate = F)+
    scale_color_manual(values=farg, na.translate = F)+
      labs(fill = "Land use", y = "Diversity")+
      #ylim(1,6)+
      theme_classic()+
  #geom_line(aes(group = Par),alpha = 0.05)+
  scale_x_discrete(limits = c("Managed","Recent","Late","trt99","Forest"),
  breaks = c("Managed","Recent","Late",NA,"Forest"),
  labels = c("Managed","Recent","Late",NA,"Forest"))+
  geom_vline(xintercept = "trt99", linetype = 2, linewidth = 0.2)+
    geom_hline(yintercept = h, linetype = 2, linewidth = 0.9, col = "black")+
  theme(axis.title.x = element_blank())+
  theme(axis.text.x = element_blank())+
    theme(axis.title.y = element_blank())+
  theme(axis.ticks.x = element_blank())+
  #theme(legend.position = "none")+
     theme(text=element_text(size=20))

#Check data distribution
hist(m1$Chitin_resp, xlab = "Response", main = "")

#Test for differences in grasslands relative to forest reference level
t1 = m1 %>% filter(Chitin_resp > 0)
mod1 <- glm(Chitin_resp ~ Succession + (1|Par)  + (1|Distance),
             family = gaussian, 
             data = t1)

#Check model disagnostics
resid <- simulateResiduals(fittedModel = mod1, plot = F)
plot(resid) #Ok

#Inspect results
summary(mod1)
summ(mod1, exp = T)

```

```{r Chitin standardized standardized}

m2 %>%
  filter(Chitin_resp_std > 0) %>%
    bind_rows(tibble(var1 = "trt99")) %>% 
  ggplot(aes(x= Succession, y = Chitin_resp, fill = Succession))+
  geom_boxplot(aes(fill= Succession),outlier.shape = NA)+
    geom_point(position=position_jitter(0.1, seed = 123),alpha=0.5)+
    scale_fill_manual(values=farg, na.translate = F)+
    scale_color_manual(values=farg, na.translate = F)+
      labs(fill = "Land use", y = "Diversity")+
      #ylim(1,6)+
      theme_classic()+
  #geom_line(aes(group = Par),alpha = 0.05)+
  scale_x_discrete(limits = c("Managed","Recent","Late","trt99","Forest"),
  breaks = c("Managed","Recent","Late",NA,"Forest"),
  labels = c("Managed","Recent","Late",NA,"Forest"))+
  geom_vline(xintercept = "trt99", linetype = 2, linewidth = 0.2)+
    geom_hline(yintercept = h, linetype = 2, linewidth = 0.9, col = "black")+
  theme(axis.title.x = element_blank())+
  theme(axis.text.x = element_blank())+
    theme(axis.title.y = element_blank())+
  theme(axis.ticks.x = element_blank())+
  #theme(legend.position = "none")+
     theme(text=element_text(size=20))

#Check data distribution
hist(m1$Chitin_resp, xlab = "Response", main = "")

#Test for differences in grasslands relative to forest reference level
t1 = m1 %>% filter(Chitin_resp_std > 0)
mod1 <- glm(Chitin_resp_std ~ Succession + (1|Par)  + (1|Distance),
             family = Gamma, 
             data = t1)

#Check model disagnostics
resid <- simulateResiduals(fittedModel = mod1, plot = F)
plot(resid) #Ok

#Inspect results
summary(mod1)
summ(mod1, exp = T)

```


                                      EFFECT SIZE PLOTS
###################################################################################################
```{r Prepare dataframe}

#Paired sites only
s1 = m1 %>% 
  count(Par) %>%
  filter(n == 2) %>%
  semi_join(m1, ., by = "Par")

#Split by land use
s1 = s1 %>% dplyr::group_split(Succession)
m = s1[[3]]
r = s1[[4]]
l = s1[[2]]
f = s1[[1]]

#Merge paired dataframes
mf = f %>% filter(Par %in% m$Par)
mf = rbind(m, mf)
mf$Succession = ordered(mf$Succession, levels = c("Managed","Forest"))
rf = f %>% filter(Par %in% r$Par)
rf = rbind(r, rf)
rf$Succession = ordered(rf$Succession, levels = c("Recent","Forest"))
lf = f %>% filter(Par %in% l$Par)
lf = rbind(l, lf)
lf$Succession = ordered(lf$Succession, levels = c("Late","Forest"))

```



```{r}
library(effectsize)

```


```{r Calculate effect sizes}

#Basal resp
v1 = hedges_g(Basal_resp ~ Succession, data = mf)
v2 = hedges_g(Basal_resp ~ Succession, data = rf)
v3 = hedges_g(Basal_resp ~ Succession, data = lf)
r1 = rbind(v1,v2,v3)
r1$Contrast = c("Managed vs Forest", "Recent vs Forest", "Late vs Forest")
r1$Substrate = "Basal_resp"
r1$Standardized = "Yes"
k1 = r1

#Glucose
v1 = hedges_g(Glucose_resp ~ Succession, data = mf)
v2 = hedges_g(Glucose_resp ~ Succession, data = rf)
v3 = hedges_g(Glucose_resp ~ Succession, data = lf)
r1 = rbind(v1,v2,v3)
r1$Contrast = c("Managed vs Forest", "Recent vs Forest", "Late vs Forest")
r1$Substrate = "Glucose_resp"
r1$Standardized = "No"
k2 = r1

#Glucose Standardized
v1 = hedges_g(Glucose_resp_std ~ Succession, data = mf)
v2 = hedges_g(Glucose_resp_std ~ Succession, data = rf)
v3 = hedges_g(Glucose_resp_std ~ Succession, data = lf)
r1 = rbind(v1,v2,v3)
r1$Contrast = c("Managed vs Forest", "Recent vs Forest", "Late vs Forest")
r1$Substrate = "Glucose_resp"
r1$Standardized = "Yes"
k3 = r1

#Glycine
v1 = hedges_g(Glycine_resp ~ Succession, data = mf)
v2 = hedges_g(Glycine_resp ~ Succession, data = rf)
v3 = hedges_g(Glycine_resp ~ Succession, data = lf)
r1 = rbind(v1,v2,v3)
r1$Contrast = c("Managed vs Forest", "Recent vs Forest", "Late vs Forest")
r1$Substrate = "Glycine_resp"
r1$Standardized = "No"
k4 = r1

#Glycine Standardized
v1 = hedges_g(Glycine_resp_std ~ Succession, data = mf)
v2 = hedges_g(Glycine_resp_std ~ Succession, data = rf)
v3 = hedges_g(Glycine_resp_std ~ Succession, data = lf)
r1 = rbind(v1,v2,v3)
r1$Contrast = c("Managed vs Forest", "Recent vs Forest", "Late vs Forest")
r1$Substrate = "Glycine_resp"
r1$Standardized = "Yes"
k5 = r1


#Oxalic acid
v1 = hedges_g(OA_resp ~ Succession, data = mf)
v2 = hedges_g(OA_resp ~ Succession, data = rf)
v3 = hedges_g(OA_resp ~ Succession, data = lf)
r1 = rbind(v1,v2,v3)
r1$Contrast = c("Managed vs Forest", "Recent vs Forest", "Late vs Forest")
r1$Substrate = "OA_resp"
r1$Standardized = "No"
k6 = r1

#Oxalic acid Standardized
v1 = hedges_g(OA_resp_std ~ Succession, data = mf)
v2 = hedges_g(OA_resp_std ~ Succession, data = rf)
v3 = hedges_g(OA_resp_std ~ Succession, data = lf)
r1 = rbind(v1,v2,v3)
r1$Contrast = c("Managed vs Forest", "Recent vs Forest", "Late vs Forest")
r1$Substrate = "OA_resp"
r1$Standardized = "Yes"
k7 = r1


#Autolyzed yeast
v1 = hedges_g(AY_resp ~ Succession, data = mf)
v2 = hedges_g(AY_resp ~ Succession, data = rf)
v3 = hedges_g(AY_resp ~ Succession, data = lf)
r1 = rbind(v1,v2,v3)
r1$Contrast = c("Managed vs Forest", "Recent vs Forest", "Late vs Forest")
r1$Substrate = "AY_resp"
r1$Standardized = "No"
k8 = r1

#Autolyzed yeast standardized
v1 = hedges_g(AY_resp_std ~ Succession, data = mf)
v2 = hedges_g(AY_resp_std ~ Succession, data = rf)
v3 = hedges_g(AY_resp_std ~ Succession, data = lf)
r1 = rbind(v1,v2,v3)
r1$Contrast = c("Managed vs Forest", "Recent vs Forest", "Late vs Forest")
r1$Substrate = "AY_resp"
r1$Standardized = "Yes"
k9 = r1


#Lignin
v1 = hedges_g(Lignin_resp ~ Succession, data = mf)
v2 = hedges_g(Lignin_resp ~ Succession, data = rf)
v3 = hedges_g(Lignin_resp ~ Succession, data = lf)
r1 = rbind(v1,v2,v3)
r1$Contrast = c("Managed vs Forest", "Recent vs Forest", "Late vs Forest")
r1$Substrate = "Lignin_resp"
r1$Standardized = "No"
k10 = r1

#Lignin standardized
v1 = hedges_g(Lignin_resp_std ~ Succession, data = mf)
v2 = hedges_g(Lignin_resp_std ~ Succession, data = rf)
v3 = hedges_g(Lignin_resp_std ~ Succession, data = lf)
r1 = rbind(v1,v2,v3)
r1$Contrast = c("Managed vs Forest", "Recent vs Forest", "Late vs Forest")
r1$Substrate = "Lignin_resp"
r1$Standardized = "Yes"
k11 = r1


#Chitin
v1 = hedges_g(Chitin_resp ~ Succession, data = mf)
v2 = hedges_g(Chitin_resp ~ Succession, data = rf)
v3 = hedges_g(Chitin_resp ~ Succession, data = lf)
r1 = rbind(v1,v2,v3)
r1$Contrast = c("Managed vs Forest", "Recent vs Forest", "Late vs Forest")
r1$Substrate = "Chitin_resp"
r1$Standardized = "No"
k12 = r1

#Chitin standardized
v1 = hedges_g(Chitin_resp_std ~ Succession, data = mf)
v2 = hedges_g(Chitin_resp_std ~ Succession, data = rf)
v3 = hedges_g(Chitin_resp_std ~ Succession, data = lf)
r1 = rbind(v1,v2,v3)
r1$Contrast = c("Managed vs Forest", "Recent vs Forest", "Late vs Forest")
r1$Substrate = "Chitin_resp"
r1$Standardized = "Yes"
k13 = r1

```


```{r Merge and write}

res = rbind(k1,k2,k3,k4,k5,k6,k7,k8,k9,k10,k11,k12,k13)

write.table(res, file = "res_effectsize.txt")

```


```{r Set plot details effect size}

farg3 = c("#bd0026","#fd8d3c","#fecc5c")

res$Contrast = ordered(res$Contrast, levels = c(
  "Managed vs Forest", "Recent vs Forest", "Late vs Forest"
))

res$Substrate = ordered(res$Substrate, levels = c(
  "Chitin_resp", "Lignin_resp", "AY_resp", "OA_resp", "Glycine_resp", "Glucose_resp", 
  "Chitin_resp_std", "Lignin_resp_std", "AY_resp_std", "OA_resp_std", 
  "Glycine_resp_std", "Glucose_resp_std", "Basal_resp"))

```


```{r Plot effect size}

#Not standardized
res %>%
  filter(Standardized == "No") %>%
  ggplot(aes(x= Substrate, y = Hedges_g, color = Contrast))+
  geom_pointrange(aes(ymin = CI_low, ymax = CI_high), position = position_dodge(0.6))+
  geom_hline(yintercept = 0, linetype = "dashed") +
  theme_classic()+
  scale_color_manual(values = farg3) +
  theme(text=element_text(size=20)) + 
  coord_flip()


#standardized
res %>%
  filter(Standardized == "Yes") %>%
  ggplot(aes(x= Substrate, y = Hedges_g, color = Contrast))+
  geom_pointrange(aes(ymin = CI_low, ymax = CI_high), position = position_dodge(0.6))+
  geom_hline(yintercept = 0, linetype = "dashed") +
  theme_classic()+
  scale_color_manual(values = farg3) +
  theme(text=element_text(size=20)) + 
  coord_flip()
  

```






